A Formalization of Complete Discrete Valuation Rings and Local Fields

### 2023-09-10 CPP_paper/CPP_paper.tex
draft section 2.1

Then, for any real number $n_\m > 1$, it is easy to check that the function
$ v_\m: K \to n_\m^{\mathbb{Z} \cup \{-\infty\}} = n_\m^\mathbb{Z} \cup \{0\}$ sending $r$ to $n_\m^{-\text{val}_v(r)}$ is a valuation on $K$, which we call the \textit{$\m$-adic valuation}. For example, when $R = \ZZ$, $K = \QQ$ and $\m = (p)$ is the ideal generated by a prime number $p$, $v_p$ is the $p$-adic (multiplicative) valuation on $\QQ$, also known as the $p$-adic absolute value.

We will represent valuations in Lean using the \href{https://leanprover-community.github.io/mathlib_docs/ring_theory/valuation/basic.html#valuation}{valuation} structure available in mathlib, which agrees with our above definition. A valuation on a ring $R$ induces a topology on $R$ for which the ring operations are continuous, hence turning $R$ into a topological ring.

Mathlib also provides the class \href{https://leanprover-community.github.io/mathlib_docs/topology/algebra/valuation.html#valued}{valued}, which bundles together a ring with a distinguished valuation and the uniform space structure induced by this valuation. This class is designed for rings in which there is a canonical valuation. For example, on the rational number $\QQ$ there are

>>>>

Then, for any real number $n_\m > 1$, it is easy to check that the function
$ v_\m: K \to n_\m^{\mathbb{Z} \cup \{-\infty\}} = n_\m^\mathbb{Z} \cup \{0\}$ sending $x$ to $n_\m^{-c_\m(x)}$ is a valuation on $K$, which we call the \textit{$\m$-adic valuation}. For example, when $R = \ZZ$, $K = \QQ$ and $\m = (p)$ is the ideal generated by a prime number $p$, $v_p$ is the $p$-adic (multiplicative) valuation on $\QQ$, also known as the $p$-adic absolute value.

Note that the choice of number $n_\m > 1$ in the definition of $v_\m$ is not relevant, in the sense that any two such choices will give rise to equivalent valuations. In our formalization, we abstract the codomain $n_\m^\mathbb{Z} \cup \{0\}$ using the type
\texttt{with\_zero (multiplicative $\mathbb{Z}$)}, denoted $\ZZ_{m0}$,
which carries the multiplicative structure corresponding to the additive structure on $\ZZ$, as well as an added zero element.

For example, the rational numbers $20$ and $25$ have $5$-adic valuation $(-1 : \ZZ_{m0})$ and $(-2 : \ZZ_{m0})$, respectively, so $25$ is smaller than $20$ in the $p$-adic topology. \mi{Do you want to add anything else about this?}

We represent valuations in Lean using the \href{https://leanprover-community.github.io/mathlib_docs/ring_theory/valuation/basic.html#valuation}{valuation} structure available in mathlib, which encodes our above definition. A valuation on a ring $R$ induces a topology on $R$ for which the ring operations are continuous, hence turning $R$ into a topological ring.

Mathlib also provides the class \href{https://leanprover-community.github.io/mathlib_docs/topology/algebra/valuation.html#valued}{valued}, which bundles together a ring with a
uniform space structure and a distinguished valuation that induces it. This class is designed for rings in which there is a canonical valuation. For example, on the field $\QQ_p$ of $p$-adic numbers, there is a unique nontrivial valuation, which we register as an instance of the valued class. By contrast, on the rational numbers there are infinitely many valuations, none of them preferred over the others, so we do not put a global valued instance on $\QQ$.

\mi{I would like to move this paragraph to section 2.2, after  valuations in DVRs have been discussed }. When implementing valuations on a discrete valuation ring $R$ with field of fractions $K$, we made the design choice to put an instance of the valued class on $K$, but not on the ring $R$ itself. The reason is that, if needed, we can recover the
topological structure on $R$ from the one on $K$, so we prefer not to duplicate this information. \mi{This choice is consistent with mathlib's file \url{ring_theory.dedekind_domain.adic_valuation}, in which valued instances are only put on fields (should we mention this?)}

Given a ring $R$ with a valuation $v$, the \textit{unit ball} \mi {or  \textit{ring of integers}} of $R$ is defined as the subring $R_0$ of elements with valuation less than or equal to one. In mathlib, this subring is denoted \href{https://leanprover-community.github.io/mathlib_docs/ring_theory/valuation/integers.html#valuation.integer}{valuation.integer}. If $R$ is a field, then $R_0$ is a \textit{valuation subring}, that is, a subring such that for every element $r$ of $R$, either $r \in R_0$ or $r^{-1} \in R_0$.
To take advantage of the results about valuation subrings available in \href{https://leanprover-community.github.io/mathlib_docs/ring_theory/valuation/valuation_subring.html#valuation_subring}{mathlib}, when $R$ is a field we use the definition \href{https://leanprover-community.github.io/mathlib_docs/ring_theory/valuation/valuation_subring.html#valuation.valuation_subring}{valuation.valuation\_subring} to define the unit ball. This should not be confused with mathlib's \href{https://leanprover-community.github.io/mathlib_docs/ring_theory/valuation/valuation_ring.html#valuation_ring}{valuation\_ring} class.
